---
title: "Logistic regression of mortality in cohort A"
author: "Stefano Giulieri"
date: "2023-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(stringr::str_c(here::here(), "/Cell_Reports_revision/ML"))
msg <- glue::glue("My directory is {getwd()}")
message(msg)
```

Here we a fit a logistic regression model of mortality in VANANZ only so that we can use both duration of bacteraemia and vancomycin MIC predictors. This analysis was requested by the Cell Reports reviewers

Logistic regression in R:
https://quantifyinghealth.com/logistic-regression-in-r-with-categorical-variables/

Logistic regression in `tidymodels`:
https://www.kaggle.com/code/j3rome/logistic-regression-tidymodels
https://medium.com/the-researchers-guide/modelling-binary-logistic-regression-using-tidymodels-library-in-r-part-1-c1bdce0ac055
https://medium.com/the-researchers-guide/modelling-binary-logistic-regression-using-tidymodels-library-in-r-part-1-c1bdce0ac055

```{r}
library(tidyverse)
library(tidymodels)
library(skimr)

rm(list = ls())
```

# General concept

We want to compare our random forest model with a more established ("standard") logistic regression model of 30-day mortality from SAB. This would allow to determine the adjusted weights (odds ratio and effect sizes, along with confidence intervals and p-values) of the predictors [host clinical, treatment and microbiology factors] on mortality.

# The data

We use the same dataset: all cohort A episodes with available predictors of interest. Also, we excluded episodes with no therapy (there was no such episode in cohort A, but there were a couple in cohort B)

```{r}
df_mortality <- read_tsv("raw_data/ML_mortality_split_v5_input_data.tab")

df_mortality %>%
  skim()
```

Better to transform character variables into factors

```{r}
df_mortality <- df_mortality %>%
  mutate(across(.cols = where(is.character),
                .fns = as.factor)) %>%
  mutate(mortality = fct_relevel(mortality, "Survived"))

df_mortality %>%
  skim()
```

# The model

```{r}
model <- logistic_reg()
# recipe <- recipe(mortality ~ ., data = df_mortality) %>%
#   step_dummy(all_factor_predictors()) 
recipe <- recipe(mortality ~ ., data = df_mortality) 
wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(model)

m1 <- wf  %>% fit(data = df_mortality)
tidy(m1, conf.int = T)
```

Print in a nicer format

```{r}
print_log_reg <- function(df){
  df <- df %>%
     filter(term != "(Intercept)") %>%
    transmute(predictor = term, or = exp(estimate), or_low = exp(conf.low), or_high = exp(conf.high), p = p.value) %>%
  arrange(p)
  
  df <- df %>%
  mutate(across(starts_with("or"),
                ~formatC(., format = "f", digits = 2))) %>%
  transmute(predictor, or = str_c(or, " (", or_low, "-", or_high, ")"), p)
  
  return(df)
}

df_multiv <- tidy(m1, conf.int = T) %>%
  print_log_reg() %>%
  rename_with(~str_c("multiv_", .), or:p)
```

# Univariate analysis

```{r}
models <- list()
for (var in colnames(df_mortality)[-1]){
  v <- rlang::sym(var)
  model <- logistic_reg() %>%
    fit(as.formula(str_c("mortality ~", var)), data = df_mortality)
  
  models[[var]] <- model
}
df_models <- tibble(predictor = names(models),
                    model = models) %>%
  group_by(predictor) %>%
  mutate(tidy = map(model, tidy, conf.int = T)) %>%
  unnest(tidy) %>%
  filter(term != "(Intercept)")
df_univ <- df_models %>%
  print_log_reg() %>%
  rename_with(~str_c("univ_", .), or:p)
df_univ
```

# Export table to the manuscript

```{r}
df_log_reg <- df_univ %>%
  left_join(df_multiv)

df_log_reg %>%
  write_csv("processed_data/logistic_regression_mortality_clinical.csv")
```

